---
title: "Watching Video Games - Diagnostics"
author: "Tiernan J. Cahill"
output: html_notebook
html_output: default
---

This script takes the cleaned output from the DATA_MERGE script and runs diagnostics and outputs univariate stats for the first part of the analysis, setting the groundwork for later hypothesis tests. 

```{r setup, include=FALSE}
library(dplyr)
library(knitr)
library(psych)
readRDS(file = "data/merged_all.rds") -> vgv
```

# Basics
First, calculate the *n* for the study based on the number of unique participant IDs in the dataset.
```{r n-calc}
vgv$id %>% unique() %>% length() -> n
```
*n* = `r n`


# Diagnostics

## Reliability

Several of the state DVs in the post-test questionnaire were multi-item batteries. While these scales have all been previously validated, we should check the reliability of our measurements nevertheless.
```{r reliability, results='asis'}
list('IMI Enjoyment'=c("IMI_Enjoyment.1", "IMI_Enjoyment.2", "IMI_Enjoyment.3.R", 
                     "IMI_Enjoyment.4.R", "IMI_Enjoyment.5", "IMI_Enjoyment.6", "IMI_Enjoyment.7"),
       'PSI Cognitive'=c("PSI_Cognitive.1", "PSI_Cognitive.2.R", "PSI_Cognitive.3",
                     "PSI_Cognitive.4", "PSI_Cognitive.5", "PSI_Cognitive.6"),
       'PSI Affective'=c("PSI_Affective.1", "PSI_Affective.2", "PSI_Affective.3.R"),
       'PSI Behavioral'=c("PSI_Behavioral.1.R", "PSI_Behavioral.2", "PSI_Behavioral.3")) %>%
  scoreItems(., vgv, impute="none") %>% 
  .$alpha %>% 
  kable(digits = 3)
```

