---
title: "Watching Video Games - Diagnostics"
author: "Tiernan J. Cahill"
output: html_notebook
html_output: default
---

This script takes the cleaned output from the DATA_MERGE script and runs diagnostics and outputs univariate stats for the first part of the analysis, setting the groundwork for later hypothesis tests. 

```{r setup, include=FALSE}
library(dplyr)
library(knitr)
library(psych)
readRDS(file = "data/merged_all.rds") -> vgv
readRDS(file = "data/Qualtrics_merged.rds") -> qualtrics
```

# Basics
First, calculate the *n* for the study based on the number of unique participant IDs in the dataset.
```{r n-calc}
vgv$id %>% unique() %>% length() -> n.all
qualtrics$id %>% unique() %>% length() -> n.qualtrics

if (n.all == n.qualtrics) {
  print(n.all)
} else {
  print("ERROR: Number of participant IDs inconsistent across datasets.")
}
```



# Diagnostics

## Reliability

Several of the state DVs in the post-test questionnaire were multi-item batteries. While these scales have all been previously validated, we should check the reliability of our measurements nevertheless.
```{r reliability, results='asis'}
list('IMI_Enjoyment'=c("IMI_Enjoyment.1", "IMI_Enjoyment.2", "IMI_Enjoyment.3.R", 
                     "IMI_Enjoyment.4.R", "IMI_Enjoyment.5", "IMI_Enjoyment.6", "IMI_Enjoyment.7"),
       'PSI_Cognitive'=c("PSI_Cognitive.1", "PSI_Cognitive.2.R", "PSI_Cognitive.3",
                     "PSI_Cognitive.4", "PSI_Cognitive.5", "PSI_Cognitive.6"),
       'PSI_Affective'=c("PSI_Affective.1", "PSI_Affective.2", "PSI_Affective.3.R"),
       'PSI_Behavioral'=c("PSI_Behavioral.1.R", "PSI_Behavioral.2", "PSI_Behavioral.3"),
       'PSI_Overall'=c("PSI_Cognitive.1", "PSI_Cognitive.2.R", "PSI_Cognitive.3",
                       "PSI_Cognitive.4", "PSI_Cognitive.5", "PSI_Cognitive.6",
                       "PSI_Affective.1", "PSI_Affective.2", "PSI_Affective.3.R",
                       "PSI_Behavioral.1.R", "PSI_Behavioral.2", "PSI_Behavioral.3"),
       'playerEfficacy' = c("playerSkill", "playerExp", "playerBetterThanAvg", "playerBetterThanViewer")) %>%
  scoreItems(., qualtrics, impute="none") %>% 
  .$alpha %>% t() %>%
  kable(col.names = "Cronbach's &alpha;", digits = 3)
```

# Univariate Statistics

## Demographics
Get some basic measures of the demographic breakdown of the sample, starting
with gender and age.
```{r demo-gender, results='asis'}
# We're going to use this again, so make it a function
freq_kable <- function(df, x, digits = 3) {
  table(df[,x]) %>% 
  rbind(prop.table(table(df[,x])) * 100) %>% 
  t() %>% kable(col.names = c("Frequency", "%"), digits = digits)
}

qualtrics %>% freq_kable("gender")
```

```{r demo-age, results='asis'}
qualtrics %>% summarise("Median Age" = median(age), IQR = IQR(age)) %>% 
  kable()
```

Since the study depends in large part on the participants being able to understand and follow along with the streamer (to the point where they can form a parasocial connection), we also want to see how many participants learned English as a first language.

```{r demo-lang, results='asis'}
qualtrics %>% freq_kable("firstLang")
```
## Measures

Get univariate measures of central tendency and variance for survey measures of interest.
```{r univar-measures, results='asis'}

vars <- c("IMI_Enjoyment", "PSI_Cognitive", "PSI_Affective", "PSI_Behavioral",
          "PSI_Overall", "playerSkill", "playerExp", "playerBetterThanAvg", 
          "playerBetterThanViewer", "playerEfficacy")
qualtrics %>% summarise_at(vars, mean, na.rm = T) -> means
qualtrics %>% summarise_at(vars, sd, na.rm = T) -> sds

bind_rows(means, sds) %>% t() %>% kable(col.names = c("Mean", "SD"), digits = 3)
```