---
title: "Watching Video Games - Analysis"
author: "Tiernan J. Cahill"
output: html_notebook
---

This script takes the cleaned output from the DATA_MERGE script and runs hypothesis tests and inductive stats for the second part of the analysis, following diagnostics and univariate stats.

```{r setup, include=FALSE}
library(dplyr)
library(tidyr)
library(knitr)
library(ggplot2)
library(ggpubr)
library(car)
library(rstatix)
readRDS(file = "data/merged_all.rds") -> vgv
readRDS(file = "data/Qualtrics_merged.rds") -> qualtrics
```


# Hypothesis Testing
## Hypothesis 1: PSI -> Enjoyment
H1 is that *viewers with a higher degree of parasocial interaction with the streamer will also experience a higher degree of enjoyment*. This can be assessed through simple bivariate correlation of the self-report measures. (*Note: In this case, we are using Spearman's correlation, because enjoyment is not normally distributed* Spearman's rank correlation coefficient makes no assumptions about the distribution of the data, unlike Pearson's product-moment correlation.)

```{r h1}
# Display a scatter plot
# ggplot(qualtrics, aes(x=PSI_Overall, y=IMI_Enjoyment)) + geom_point() +
#  xlab("PSI") + ylab("Enjoyment")

# Display a plot of rank correlation (reflecting Sperman's coefficient)
qualtrics %>% 
  transmute(PSI_Overall = rank(PSI_Overall),
            IMI_Enjoyment = rank(IMI_Enjoyment)) %>%
  ggplot(aes(x=PSI_Overall, y=IMI_Enjoyment)) + geom_point() +
    geom_smooth(method=lm, se=F) +
    xlab("PSI (by rank)") + ylab("Enjoyment (by rank)")

cor.test(~ PSI_Overall + IMI_Enjoyment, qualtrics, alternative = "greater", method="spearman") -> h1.test # One-sided test
print(h1.test)
```

The hypothesis is **supported**: *&rho;* = `r h1.test$estimate %>% round(3)`, *p* = `r h1.test$p.value %>% round(3)`. **(Note for future reference: all reports of *p = 0* are actually *p < .001* due to Base R's rounding function.)**

## Hypothesis 2: Facecam -> PSI
H2 is that *viewers in the condition where the facecam was present will experience a higher degree of parasocial interaction than viewers in the condition where the facecam was absent*. This can be assessed with an independent-sample T test.

```{r h2}
# First, we need a version of the full dataset with one row per participant, 
# containing their assigned condition and their self-reported PSI
vgv %>% select(id, stimFace, PSI_Overall) %>% unique() -> h2.data

# Visualize the difference between groups using a boxplot
ggboxplot(h2.data, x = "stimFace", y = "PSI_Overall",
          xlab = "Facecam Present", ylab = "PSI")

# First test for homogeneity of variance
bartlett.test(PSI_Overall ~ stimFace, h2.data) -> h2.bartlett_test
h2.var_equal <- if(h2.bartlett_test$p.value > 0.05) TRUE else FALSE
print(h2.bartlett_test)

t.test(PSI_Overall ~ stimFace, h2.data, var.equal = h2.var_equal, 
       alternative = "less") -> h2.test # One-sided test
print(h2.test)
```

The hypothesis is **supported**: *t*(`r h2.test$parameter`) = `r h2.test$statistic %>% round(3)`, *p* = `r h2.test$p.value %>% round(3)`. Note that `r h2.data %>% summarise(sum(is.na(PSI_Overall))) %>% .[[1]]` cases that did not complete the PSI questionnaire battery were removed from this analysis.

## Hypothesis 3: Streamer Gender -> Perceived Efficacy
H3 is that *the gender of the streamer will affect the perceived efficacy of the streamer*, given identical gameplay performance. This can also be assessed with a Mann-Whitney test of stochastic equality (since perceived efficacy is not normally distributed).

```{r h3}
# First, we need a version of the full dataset with one row per participant, 
# containing their assigned condition and their perception of the streamer's
# efficacy
vgv %>% select(id, stimGender, playerEfficacy) %>% unique() -> h3.data

# Visualize the difference between groups using a boxplot
ggboxplot(h3.data, x = "stimGender", y = "playerEfficacy",
          color = "stimGender", legend = "none",
          xlab = "Streamer Gender", ylab = "Perceived Efficacy")

# It appears that there are some outliers on the low end of perceived efficacy 
# in both groups; these should be removed before running tests
h3.outliers <- boxplot(playerEfficacy ~ stimGender, data = h3.data, plot=F)$out
h3.data <- h3.data[-which(h3.data$playerEfficacy %in% h3.outliers),]

wilcox.test(playerEfficacy ~ stimGender, h3.data, 
       alternative = "two.sided", correct=F) -> h3.test # Two-sided test
print(h3.test)
```

The hypothesis is **not supported**: *U* = `r h3.test$statistic %>% round(3)`, *p* = `r h3.test$p.value %>% round(3)`. Note that `r length(h3.outliers)` outliers were removed from the analysis.

### Control for Experience
Another possibility is that audience's perception of player skill is moderated by their own experience playing games. The following analysis investigates the hypothesized relationship from H3 while controlling for experience *playing games* and for experience *watching video game content online*.

```{r h3b}
# In this case, we want a version of the dataset that includes the previous 
# variables from H3, but with the addition of measures of frequency of play
vgv %>% select(id, stimGender, playerEfficacy, 
               hrsWatchingRecordedVGContent_weekday,
               hrsWatchingRecordedVGContent_weekend,
               hrsWatchingStreamedVGContent_weekday,
               hrsWatchingStreamedVGContent_weekend,
               hrsPlayingSinglePlayer_weekday,
               hrsPlayingSinglePlayer_weekend,
               hrsPlayingMultiPlayer_weekday,
               hrsPlayingMultiPlayer_weekend) %>% unique() -> h3b.data

# Recode non-responses as zero hours
h3b.data[is.na(h3b.data)] = 0

# We can combine some of these measures together--since this level of granularity
# probably isn't particularly helpful--to get aggregate levels of gameplay
# and viewership per week
h3b.data %>% transmute(id = id,
                       stimGender = stimGender,
                       playerEfficacy = playerEfficacy,
                       hrsWatchingVGContent = hrsWatchingRecordedVGContent_weekday * 5 + 
                                              hrsWatchingRecordedVGContent_weekend * 2 +
                                              hrsWatchingStreamedVGContent_weekday * 5 +
                                              hrsWatchingStreamedVGContent_weekend * 2,
                       hrsPlayingGames = hrsPlayingSinglePlayer_weekday * 5 +
                                         hrsPlayingSinglePlayer_weekend * 2 +
                                         hrsPlayingMultiPlayer_weekday * 5 +
                                         hrsPlayingMultiPlayer_weekend * 2) -> h3b.data



# Let's visualize the results
h3b.data %>% ggplot(aes(x=x)) +
  geom_histogram(aes(x = hrsPlayingGames), fill = "#69b3a2", alpha = 0.4, binwidth = 5) +
  geom_label(aes(x = 75, y = 35, label = "Hrs Playing Games"), color = "#69b3a2") +
  geom_histogram(aes(x = hrsWatchingVGContent), fill = "#404080", alpha = 0.4, binwidth = 5) +
  geom_label(aes(x = 75, y = 25, label = "Hrs Watching Games"), color = "#404080")

# Remove outliers 
h3b.data %>% filter(hrsPlayingGames < 8*7 & hrsWatchingVGContent < 8*7) -> h3b.data
# h3b.play_outliers <- boxplot(h3b.data$hrsPlayingGames, plot=F)$out
# h3b.watch_outliers <- boxplot(h3b.data$hrsWatchingVGContent, plot=F)$out
# h3b.data <- h3b.data[-which(h3b.data$hrsPlayingGames %in% h3b.play_outliers),]
#h3b.data <- h3b.data[-which(h3b.data$hrsWatchingVGContent %in% h3b.watch_outliers),]

# Evaluate correlation between potential confounders (two measures of experience)
# and the outcome variable (player efficacy)
cor.test(~ playerEfficacy + hrsPlayingGames, h3b.data)
cor.test(~ playerEfficacy + hrsWatchingVGContent, h3b.data)

# Only times spent watching VG content online seems to correlate with perceived
# player efficacy, so this should be included in the model as a covariate

ggscatter(h3b.data, x = "hrsWatchingVGContent", y = "playerEfficacy",
          color = "stimGender", add = "reg.line")

# Check the assumption of homogeneity of regression slopes
h3b.data %>% anova_test(playerEfficacy ~ stimGender*hrsWatchingVGContent)

# Check the assumption of normality of residuals
h3b.model <- lm(playerEfficacy ~ hrsWatchingVGContent + stimGender, h3b.data)
h3b.model.metrics <- augment(h3b.model) %>% 
  select(-.hat, -.sigma, -.fitted, -.se.fit)
shapiro_test(h3b.model.metrics$.resid)

# It appears that there are more outliers as identified using the standardized 
# residual -- these outliers seem to occur when both the outcome and the 
# covariate are equal to 0
h3b.model.metrics %>% filter(abs(.std.resid) > 3) %>% as.data.frame()
h3b.data <- h3b.data %>%filter(!(playerEfficacy == 0 & hrsWatchingVGContent == 0))

# Refit the model
h3b.model <- lm(playerEfficacy ~ hrsWatchingVGContent + stimGender, h3b.data)
h3b.model.metrics <- augment(h3b.model) %>% 
  select(-.hat, -.sigma, -.fitted, -.se.fit)

# Confirm homogeneity of variance of residuals
bartlett.test(.resid ~ stimGender, h3b.model.metrics)

# Fit and display ANCOVA model
h3b.model <- aov(playerEfficacy ~ hrsWatchingVGContent + stimGender, h3b.data)
Anova(h3b.model, type = "III")
```
Visualization suggests that most respondents don't play games, but do watch some, and that there are a small number of extreme outliers in terms of both playing and watching (i.e., the figures are right-skewed). These outliers (*n* = `r nrow(qualtrics) - nrow(h3b.data)`) have been removed from the ANCOVA.

The ANCOVA model suggests that while there are still not significant differences in perceived player efficacy as a result of the streamer's gender, even when controlling for the potential confounding factor of experience.

```{r h3b_p2}
# Also check if there's a significant difference between players and non-players,
# watchers and non-watchers
h3b.data %>% mutate(player = hrsPlayingGames > 0,
                    watcher = hrsWatchingVGContent > 0) -> h3b.data

wilcox.test(playerEfficacy ~ player, h3b.data, 
       alternative = "two.sided", correct = FALSE)

wilcox.test(playerEfficacy ~ watcher, h3b.data,
            alternative = "two.sided", correct = FALSE)
```

While the hypothesis is **not supported**, this seems to suggest that time spent
watching VG content is more influential in determining perceived efficacy than either
gender or play experience.

```{r h3_exp}
h3_exp.model <- lm(playerEfficacy ~ hrsWatchingVGContent, h3b.data)
summary(h3_exp.model)

# This model has some issues -- come back and respecify
par(mfrow = c(2, 2))
plot(h3_exp.model)

h3b.data %>% ggscatter(x="hrsWatchingVGContent", y = "playerEfficacy", color="stimGender") + stat_smooth(method = "lm")
```
Interestingly, this suggests a small yet significant positive effect of past viewership on the perceived efficacy of the streamer.

## Hypothesis 4: Attention -> Parasocial Interaction
H4 is that *viewers who allocate more attention to the streamer's face (where the facecam is displayed) will experience a proportionately greater level of parasocial interaction*.
```{r h4}
# First, create a version of the dataset with one row per participant, 
# containing their self-reported level of PSI, as well as an aggregate
# measure of their relative attention allocation to the facecam (thus containg
# participants in the facecam condition only )

# vgv %>% 
#   filter(aoi == "FACE") %>%
#   group_by(id) %>%
#   summarise(faceRatioG = mean(perSpentG),
#             faceRatioF = mean(perSpentF),
#             PSI_Overall = first(PSI_Overall)) ->
#   h4.data

# Alternate method of calculating attention allocation
vgv %>%
  filter(stimFace == TRUE) %>%
  group_by(id, aoi) %>%
  summarise(totalG = sum(timeSpentG),
            totalF = sum(timeSpentF),
            PSI_Overall = first(PSI_Overall)) %>%
  ungroup() %>%
  arrange(id, aoi) %>%
  mutate(faceRatioG = totalG / (totalG+lead(totalG)),
         faceRatioF = totalF / (totalF+lead(totalF))) %>%
  filter(aoi == "FACE") %>%
  select(-aoi, -totalG, -totalF) -> h4.data

# Generate a scatter plot for gaze ratio
ggplot(h4.data, aes(x=faceRatioG, y=PSI_Overall)) + geom_point() +
  geom_smooth(method = lm) +
  xlab("Ratio of Gaze on Facecam") + ylab("PSI")

# Generate a scatter plot for fixation ratio
ggplot(h4.data, aes(x=faceRatioF, y=PSI_Overall)) + geom_point() +
  geom_smooth(method = lm) +
  xlab("Ratio of Fixation on Facecam") + ylab("PSI")

# Test the correlation for gaze using Spearman because gaze ratio
# isn't normally distributed)
cor.test(~ PSI_Overall + faceRatioG, h4.data,
         alternative = "greater",
         method="spearman") -> h4.test.g
print(h4.test.g)

# Test the correlation for fixation (
cor.test(~ PSI_Overall + faceRatioF, h4.data,
         alternative = "greater") -> h4.test.f
print(h4.test.f)
```
The hypothesis is **not supported** for gaze measurement: *&rho;* = `r h4.test.g$estimate %>% round(3)`, *p* = `r h4.test.g$p.value %>% round(3)`.

The hypothesis is **not supported** for fixation measurement: *&rho;* = `r h4.test.f$estimate %>% round(3)`, *p* = `r h4.test.f$p.value %>% round(3)`.


## Hypothesis 5: Attention -> Arousal
H5 is that *viewers who allocate more attention to the streamer's face (where the facecam is displayed) will experience a proportionately higher level of emotional arousal*.
```{r h5}
# First, create a version of the dataset with one row per participant, 
# containing aggregate measures of their relative attention allocation to the
# facecam (thus containing participants in the facecam condition only ) and 
# their arousal (GSR)

vgv %>%
  filter(stimFace == TRUE) %>%
  group_by(id, aoi) %>%
  summarise(totalG = sum(timeSpentG),
            totalF = sum(timeSpentF),
            peakCount = sum(peakCount),
            peaksPerMin = mean(peaksPerMin)) %>%
  ungroup() %>%
  arrange(id, aoi) %>%
  mutate(faceRatioG = totalG / (totalG+lead(totalG)),
         faceRatioF = totalF / (totalF+lead(totalF))) %>%
  filter(aoi == "FACE") %>%
  select(-aoi, -totalG, -totalF) -> h5.data

# There are a few outliers in the arousal data, so use the simple 2*IQR rule 
# to filter them out.
h5.data.no_outliers <- h5.data %>% 
  filter(peaksPerMin < median(h5.data$peaksPerMin, na.rm=T) + 2*IQR(h5.data$peaksPerMin, na.rm=T)) %>%
  filter(peaksPerMin > median(h5.data$peaksPerMin, na.rm=T) -
           2*IQR(h5.data$peaksPerMin, na.rm=T))

# Generate a scatter plot for gaze ratio
ggplot(h5.data.no_outliers, aes(x=faceRatioG, y=peaksPerMin)) + geom_point() +
  geom_smooth(method = lm) +
  xlab("Ratio of Gaze on Facecam") + ylab("Arousal")

# Generate a scatter plot for fixation ratio
ggplot(h5.data.no_outliers, aes(x=faceRatioF, y=peaksPerMin)) + geom_point() +
  geom_smooth(method = lm) +
  xlab("Ratio of Fixation on Facecam") + ylab("Arousal")

# Test the correlation for gaze (using Spearman because gaze ratio
# isn't normally distributed)
cor.test(~ peaksPerMin + faceRatioG, h5.data.no_outliers,
         alternative = "greater",
         method="spearman") -> h5.test.g
print(h5.test.g)

# Test the correlation for fixation (using Spearman because GSR data isn't
# normally distributed)
cor.test(~ peaksPerMin + faceRatioF, h5.data.no_outliers,
         alternative = "greater",
         method="spearman") -> h5.test.f
print(h5.test.f)
```
The hypothesis is **not supported** for gaze measurement: *&rho;* = `r h5.test.g$estimate %>% round(3)`, *p* = `r h5.test.g$p.value %>% round(3)`.

The hypothesis is **not supported** for fixation measurement: *&rho;* = `r h5.test.f$estimate %>% round(3)`, *p* = `r h5.test.f$p.value %>% round(3)`.

## Hypothesis 6: Arousal -> Enjoyment
H6 is that *viewers who experience a higher level of emotional arousal will also experience a proportionately higher degree of enjoyment*.

```{r h6}
# First, create a version of the dataset with one row per participant, 
# containing their self-reported level of enjoyment, as well as an aggregate
# measure of their arousal (GSR) across both stimuli
vgv %>% filter(aoi == "GAME") %>%
  group_by(id) %>% 
  summarise(peakCount = sum(peakCount), 
            peaksPerMin = mean(peaksPerMin), 
            IMI_Enjoyment = first(IMI_Enjoyment)) -> 
  h6.data

# There are a few outliers in the biometric data, so use the simple 2*IQR rule 
# to filter them out.
h6.data.no_outliers <- h6.data %>% 
  filter(peaksPerMin < median(h6.data$peaksPerMin, na.rm=T) +
           2*IQR(h6.data$peaksPerMin, na.rm=T)) %>%
  filter(peaksPerMin > median(h6.data$peaksPerMin, na.rm=T) -
           2*IQR(h6.data$peaksPerMin, na.rm=T))

# Generate a scatter plot 
ggplot(h6.data.no_outliers, aes(x=peaksPerMin, y=IMI_Enjoyment)) + geom_point() +
  geom_smooth(method = lm) +
  xlab("Arousal") + ylab("Enjoyment")

# Test the correlation (using Spearman again, since the variables are not
# normally distributed)
cor.test(~ IMI_Enjoyment + peaksPerMin, h6.data.no_outliers, 
         alternative = "greater",
         method="spearman") -> h6.test
print(h6.test)
```

The hypothesis is **not supported**: *&rho;* = `r h6.test$estimate %>% round(3)`, *p* = `r h6.test$p.value %>% round(3)`.