---
title: "Watching Video Games - Analysis"
author: "Tiernan J. Cahill"
output: html_notebook
---

This script takes the cleaned output from the DATA_MERGE script and runs hypothesis tests and inductive stats for the second part of the analysis, following diagnostics and univariate states.

```{r setup, include=FALSE}
library(dplyr)
library(knitr)
readRDS(file = "data/merged_all.rds") -> vgv
readRDS(file = "data/Qualtrics_merged.rds") -> qualtrics
```


# Hypothesis Testing
## Hypothesis 1: PSI -> Enjoyment
H1 is that *viewers with a higher degree of parasocial interaction with the streamer will also experience a higher degree of enjoyment*. This can be assessed through simple bivariate correlation of the self-report measures.

```{r h1}
cor.test(~ PSI_Overall + IMI_Enjoyment, qualtrics) -> h1.test
print(h1.test)
```

The hypothesis is **supported** (*r* = `r h1.test$estimate %>% round(3)`, *p* = `r h1.test$p.value %>% round(3)`). **(Note for future reference: all reports of *p = 0* are actually *p < .001* due to Base R's rounding function.)**