---
title: "Watching Video Games - Secondary Analysis"
author: "Tiernan J. Cahill"
output: html_notebook
---

This script takes the cleaned output from the DATA_MERGE script and runs some exploratory data analysis, following hypothesis testing, to look at some of our additional research questions.

```{r setup, include=FALSE}
library(dplyr)
library(knitr)
readRDS(file = "data/merged_all.rds") -> vgv
readRDS(file = "data/Qualtrics_merged.rds") -> qualtrics
```

# Research Questions
## RQ1: Gender -> PSI
RQ1 is: *Will viewers whose sexual orientation aligns with the gender of the streamer experience different levels of parasocial interaction (PSI) than others?*
```{r rq1}
```

## RQ2: Gender -> Arousal
RQ2 is: *Will viewers whose sexual orientation aligns with the gender of the streamer experience different levels of physiological arousal than others?*
```{r rq2}
```

## RQ3: Gender -> Attention
RQ3 is: *Will viewers whose sexual orientation aligns with the gender of the streamer allocate their attention differently than others?*
```{r rq3}
```