---
title: "Watching Video Games - Secondary Analysis"
author: "Tiernan J. Cahill"
output: html_notebook
---

This script takes the cleaned output from the DATA_MERGE script and runs some exploratory data analysis, following hypothesis testing, to look at some of our additional research questions.

```{r setup, include=FALSE}
library(dplyr)
library(knitr)
readRDS(file = "data/merged_all.rds") -> vgv
readRDS(file = "data/Qualtrics_merged.rds") -> qualtrics
```

# Research Questions
## RQ1: Gender -> PSI
RQ1 is: *Will viewers whose sexual orientation aligns with the gender of the streamer experience different levels of parasocial interaction (PSI) than others?*

This can be answered looking at self-report data alone, but first we need to determine for each participant whether their sexual orientation aligns with the gender of the streamer (i.e., for a participant who reports that they are interested in women or in all genders and were assigned to the condition with the female streamer).
```{r calc-orientation, results='asis'}
vgv <- vgv %>% 
  mutate(orientationMatch = ((stimGender == "FEMALE" & orientation == "Women") |
                             (stimGender == "MALE" & orientation == "Men") |
                             (orientation == "Both")))

# Let's take a quick look at the breakdown of orientations and matches
# Recrate the function from the beginning of the analysis
freq_kable <- function(df, x, digits = 3) {
  table(df[,x]) %>% 
  rbind(prop.table(table(df[,x])) * 100) %>% 
  t() %>% kable(col.names = c("Frequency", "%"), digits = digits)
}

vgv_unique <- vgv %>% 
  select(id, stimGender, gender, orientation, orientationMatch) %>%
  unique()

vgv_unique %>% freq_kable("orientation")
vgv_unique %>% freq_kable("orientationMatch")
```
Note that `r vgv_unique %>% filter(is.na(orientation)) %>% count() %>% .[[1]]` participants did not answer the question about sexual orientation and were excluded from subsequent analysis.

```{r rq1}
```

## RQ2: Gender -> Arousal
RQ2 is: *Will viewers whose sexual orientation aligns with the gender of the streamer experience different levels of physiological arousal than others?*
```{r rq2}
```

## RQ3: Gender -> Attention
RQ3 is: *Will viewers whose sexual orientation aligns with the gender of the streamer allocate their attention differently than others?*
```{r rq3}
```