---
title: "Watching Video Games - Secondary Analysis"
author: "Tiernan J. Cahill"
output: html_notebook
---

This script takes the cleaned output from the DATA_MERGE script and runs some exploratory data analysis, following hypothesis testing, to look at some of our additional research questions.

```{r setup, include=FALSE}
library(dplyr)
library(knitr)
library(ggplot2)
library(ggpubr)
readRDS(file = "data/merged_all.rds") -> vgv
readRDS(file = "data/Qualtrics_merged.rds") -> qualtrics
```

# Research Questions
## RQ1: Gender -> PSI
RQ1 is: *Will viewers whose sexual orientation aligns with the gender of the streamer experience different levels of parasocial interaction (PSI) than others?*

This can be answered looking at self-report data alone, but first we need to determine for each participant whether their sexual orientation aligns with the gender of the streamer (i.e., for a participant who reports that they are interested in women or in all genders and were assigned to the condition with the female streamer).
```{r calc-orientation, results='asis'}
vgv <- vgv %>% 
  mutate(orientationMatch = ((stimGender == "FEMALE" & orientation == "Women") |
                             (stimGender == "MALE" & orientation == "Men") |
                             (orientation == "Both")))

# Let's take a quick look at the breakdown of orientations and matches
# Recrate the function from the beginning of the analysis
freq_kable <- function(df, x, digits = 3) {
  table(df[,x]) %>% 
  rbind(prop.table(table(df[,x])) * 100) %>% 
  t() %>% kable(col.names = c("Frequency", "%"), digits = digits)
}

vgv_unique <- vgv %>% 
  select(id, stimGender, gender, orientation, orientationMatch) %>%
  unique()

vgv_unique %>% freq_kable("orientation")
vgv_unique %>% freq_kable("orientationMatch")
```
Note that `r vgv_unique %>% filter(is.na(orientation)) %>% count() %>% .[[1]]` participants did not answer the question about sexual orientation and were excluded from subsequent analysis.

```{r rq1}
rq1.data <- vgv %>% 
  select(id, orientationMatch, PSI_Overall) %>%
  filter(!is.na(orientationMatch)) %>%
  unique()

# Visualize the difference between groups using a boxplot
ggboxplot(rq1.data, x = "orientationMatch", y = "PSI_Overall",
          xlab = "Orientation Match", ylab = "PSI",
          color = "orientationMatch", legend = "none")

# First test for homogeneity of variance
bartlett.test(PSI_Overall ~ orientationMatch, rq1.data)

# Run a t-test
t.test(PSI_Overall ~ orientationMatch, data = rq1.data,
       var.equal = TRUE, alternative = "two.sided")
```

From these results, it doesn't appear that there is a meaningful difference in 
parasocial interaction between the two groups of interest.

## RQ2: Gender -> Arousal
RQ2 is: *Will viewers whose sexual orientation aligns with the gender of the streamer experience different levels of physiological arousal than others?*
```{r rq2}
# First, create a version of the dataset with one row per participant,
# an aggregate measure of their arousal (GSR) across both stimuli
rq2.data <- vgv %>% 
  filter(aoi == "GAME") %>%
  group_by(id) %>% 
  summarise(peakCount = sum(peakCount), 
            peaksPerMin = mean(peaksPerMin),
            orientationMatch = first(orientationMatch)) %>%
  filter(!is.na(orientationMatch))

# Visualize the difference between groups using a boxplot
ggboxplot(rq2.data, x = "orientationMatch", y = "peaksPerMin",
          xlab = "Orientation Match", ylab = "GSR",
          color = "orientationMatch", legend = "none")

# There are a few outliers in the biometric data, so use the simple 2*IQR rule 
# to filter them out.
rq2.data.no_outliers <- rq2.data %>% 
  filter(peaksPerMin < median(rq2.data$peaksPerMin, na.rm=T) +
           2*IQR(rq2.data$peaksPerMin, na.rm=T)) %>%
  filter(peaksPerMin > median(rq2.data$peaksPerMin, na.rm=T) -
           2*IQR(rq2.data$peaksPerMin, na.rm=T))

# Visualize again
ggboxplot(rq2.data.no_outliers, x = "orientationMatch", y = "peaksPerMin",
          xlab = "Orientation Match", ylab = "GSR",
          color = "orientationMatch", legend = "none")

# Now test for homogeneity of variance
bartlett.test(peaksPerMin ~ orientationMatch, rq2.data.no_outliers)

# Run a t-test
t.test(peaksPerMin ~ orientationMatch, data = rq2.data.no_outliers,
       var.equal = TRUE, alternative = "two.sided")
```
From these results, it also doesn't appear that there is a meaningful difference in 
physiological arousal between the two groups of interest.

## RQ3: Gender -> Attention
RQ3 is: *Will viewers whose sexual orientation aligns with the gender of the streamer allocate their attention differently than others?*
```{r rq3}
# First, create a version of the dataset with one row per participant, 
# containing aggregate measures of their relative attention allocation to the
# facecam (thus containing participants in the facecam condition only)

rq3.data <- vgv %>%
  filter(stimFace == TRUE) %>%
  group_by(id, aoi) %>%
  summarise(totalG = sum(timeSpentG),
            totalF = sum(timeSpentF),
            orientationMatch = first(orientationMatch)) %>%
  ungroup() %>%
  arrange(id, aoi) %>%
  mutate(faceRatioG = totalG / (totalG+lead(totalG)),
         faceRatioF = totalF / (totalF+lead(totalF))) %>%
  filter(aoi == "FACE") %>%
  select(-aoi, -totalG, -totalF) %>%
  filter(!is.na(orientationMatch))

# Visualize the difference between groups using a boxplot
ggboxplot(rq3.data, x = "orientationMatch", y = "faceRatioG",
          xlab = "Orientation Match", ylab = "Gaze Ratio",
          color = "orientationMatch", legend = "none")

# Visualize the difference between groups using a boxplot
ggboxplot(rq3.data, x = "orientationMatch", y = "faceRatioF",
          xlab = "Orientation Match", ylab = "Fixation Ratio",
          color = "orientationMatch", legend = "none")

# Now test for homogeneity of variance
bartlett.test(faceRatioG ~ orientationMatch, rq3.data)
bartlett.test(faceRatioF ~ orientationMatch, rq3.data)

# Run some t-tests
t.test(faceRatioG ~ orientationMatch, data = rq3.data,
       var.equal = TRUE, alternative = "two.sided")

t.test(faceRatioF ~ orientationMatch, data = rq3.data,
       var.equal = TRUE, alternative = "two.sided")
```

From these results, it appears that there may be a slightly higher allocation of attention to the facecam among participants whose orientation matched the gender of the streamer, but the difference was marginal and did not approach statistical significance.